Low Complexity Adaptive Sonar Imaging

### 2015-12-17 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:12:17	09:57:16	(to.17.des.)

\newif\ifBuildBibliography\BuildBibliographytrue  % Whether to include the references.bib file or use inline references.

@@

LCA seeks to improve image resolution and contrast by minimizing the power of noise and interference in each image pixel. This is the same optimization criterion as MVDR uses, but instead of computing an optimal array window like MVDR does, it selects the best window out of a predefined set. Hence, it can be viewed as either a reduced and discrete window space version of MVDR, or as an advanced multi-apodization technique. What we have studied in this paper is which windows we should ideally let LCA choose from.

We have found the LCA to be surprisingly insensitive to the exact window function it is used with, supported by the observation that near identical images were produced whether we used Kaiser or trigonometric windows. However, it needs to be able to choose between wide and narrow window responses, since the wide ones offer the best sidelobe suppression and the narrow ones offer the best sensitivity. We achieved good results using 2-3 Kaiser windows in the range $\beta\in[0,5]$, and observed minimal difference when adding more windows or adjusting the upper bound of $\beta$. For the steering angle we suggest a value in the range $\phi\in[0,\pm\delta\phi_\text{3dB}(\beta)]$, where $\phi_\text{3dB}(\beta)$ is the bandwidth of the respective window, and $\delta$ is a scaling parameters. Setting $\delta$ to either 50\% or 100\% reduces the lateral point target size to 73\% or 53\% compared to the rectangular window, respectively. Hence, this parameter can be used to control the aggressiveness and resolution gain of LCA.

In summary, LCA is an attractive alternative to other adaptive beamformers due to being very fast, simple to understand, practically parameter-free, inherently robust, and able produce images similar to that of MVDR. Contrary to MVDR, it has a low computational complexity and can be easily and efficiently accelerated using e.g. GPUs. Since LCA performs just as well if we used trigonometric windows instead of Kaiser, we can even compute the optimal window parameter analytically for any given steering angle, and avoid the search over the window parameter entirely.

@@

Throughout this article we have used LCA with the Kaiser window function. However, as shown in \Fig{image_quality} we can obtain similar performance using the trigonometric window function. For this window we can obtain an analytic solution for the optimal value of $\alpha$ in the minimum variance sense. The derivation generalizes the one described for a non-steered window in~\cite{Stoica2005} to also apply for the steered (complex) window. We start by inserting the trigonometric function (\ref{eq:trig_window_function}), with applied steering (\ref{eq:steering_vector}), into the beamformer equation (\ref{eq:beamformer_output}):
%
\begin{align}
\w\H\x &= \sumb{m=0}{M-1} [\a_\phi]_m\Big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\Big) x_m^* \nn
&= \alpha \a_\phi\T\x - (1-\alpha)\b_\phi\T\x
\end{align}
%
where $[\a_\phi]_m$ is the $m$th component of the steering vector $\a_\phi$ defined in (\ref{eq:steering_vector}), and 
%
\begin{align}
\b_\phi = \text{diag}(\a_\phi)\cdot\begin{bmatrix}
     1 &
     \cos\left(\frac{2\pi}{M-1}\right) &
     \cos\left(\frac{2\pi 2}{M-1}\right) &
     \hdots &
     1
     \end{bmatrix}\T.
\end{align}
%
Unity gain in the look direction is ensured as long as the weights sum to one:
%
\begin{align}
\w\T\1 &= \big(\alpha\a_\phi - (1-\alpha)\b_\phi\big)\T\1 = 1.
\end{align}
%
This is true for any value of $\alpha$ if $\a_\phi\T\1 = 1$ and $\b_\phi\T\1 = 1$. Hence, we  preserve the unity gain constraint as long as we normalize $\a_\phi$ and $\b_\phi$.

@@

@@

>>>>

\newif\ifBuildBibliography\BuildBibliographyfalse  % Whether to include the references.bib file or use inline references.

@@

LCA seeks to improve image resolution and contrast by minimizing the power of noise and interference in each image pixel. This is the same optimization criterion as MVDR uses, but instead of computing an optimal array window like MVDR does, it selects the best window out of a predefined set. Hence, it can be viewed as either a reduced and discrete window space version of MVDR, or as an advanced multi-apodization technique. In this paper we have studied which windows we should ideally let LCA choose from.

We found the LCA to be surprisingly insensitive to the exact window function it is used with, supported by the observation that near identical images were produced whether we used Kaiser windows or trigonometric ones. However, LCA must be able to choose between wide and narrow window responses, since the wide ones offer the best sidelobe suppression and the narrow ones offer the best sensitivity. We achieved good results with 2-3 Kaiser windows in the range $\beta\in[0,5]$, and observed minimal difference when adding more windows or adjusting the upper bound of $\beta$. For the steering angle we suggest a value in the range $\phi\in[\pm\delta\phi_\text{3dB}(\beta)]$, where $\phi_\text{3dB}(\beta)$ is the bandwidth of the respective window, and $\delta$ is a scaling parameter. Setting $\delta$ to either 50\% or 100\% reduces the lateral point target size to 73\% or 53\% compared to the rectangular window, respectively. Hence, this parameter can be used to control the aggressiveness and resolution gain of LCA.

In summary, LCA is an attractive alternative to other adaptive beamformers due to being very fast, simple to understand, practically parameter-free, inherently robust, and able produce images similar to that of MVDR. Contrary to MVDR, it has a low computational complexity and can be easily and efficiently accelerated using e.g. GPUs. Since LCA performs just as well if we use trigonometric windows instead of Kaiser ones, we can also compute the optimal window parameter analytically for any given steering angle, and avoid the search over the window parameter entirely.

@@

Throughout this article we have used LCA with the Kaiser window function. However, as shown in \Fig{image_quality} we can obtain similar performance using the trigonometric window function. For this window we can obtain an analytic solution for the optimal value of $\alpha$ in the minimum variance sense. The following derivation generalizes that of the non-steered (real) window described in e.g.~\cite{Stoica2005} to also apply for the steered (complex) window. We start by inserting the trigonometric function (\ref{eq:trig_window_function}), with applied steering (\ref{eq:steering_vector}), into the beamformer equation (\ref{eq:beamformer_output}):
%
\begin{align}
\w\H\x &= \sumb{m=0}{M-1} [\a_\phi]_m\Big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\Big) x_m^* \nn
&= \alpha \a_\phi\T\x - (1-\alpha)\b_\phi\T\x
\end{align}
%
where $[\a_\phi]_m$ is the $m$th component of the steering vector $\a_\phi$ defined in (\ref{eq:steering_vector}), and 
%
\begin{align}
\b_\phi = \text{diag}(\a_\phi)\cdot\begin{bmatrix}
     1 &
     \cos\left(\frac{2\pi}{M-1}\right) &
     \cos\left(\frac{2\pi 2}{M-1}\right) &
     \hdots &
     1
     \end{bmatrix}\T.
\end{align}
%
Unity gain in the look direction is ensured as long as the weights sum to one:
%
\begin{align}
\w\T\1 &= \big(\alpha\a_\phi - (1-\alpha)\b_\phi\big)\T\1 = 1.
\end{align}
%
This is true for any value of $\alpha$ if $\a_\phi\T\1 = 1$ and $\b_\phi\T\1 = 1$. Hence, we  preserve the unity gain constraint as long as we normalize $\a_\phi$ and $\b_\phi$.

@@

\bibitem{Buskenes2011}
J.~I. Buskenes, C.-I.~C. Nilsen, and A.~Austeng, ``{A Low Complexity Adaptive
  Beamformer for Active Sonar Imaging},'' in \emph{Proceedings of the
  Underwater Acoustic Measurements: Technologies \& Results}.\hskip 1em plus
  0.5em minus 0.4em\relax Underwater Acoustic Measurements: Technologies \&
  Results, 2011.

@@

\bibitem{Stoica2005}
P.~Stoica and R.~L. Moses, ``{Nonparametric Methods},'' in \emph{Spectral
  Analysis of Signals}.\hskip 1em plus 0.5em minus 0.4em\relax Prentice Hall,
  2005, ch.~2, pp. 23--89.
